Study 1 Analysis

### 2024-10-23 study1.qmd
updates

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

@@

s1 <- s1 |> group_by(id) |> 
  mutate(trial=row_number(), block=ifelse(trial>=12,2,1), plan=as_factor(plan)) |> 
  select(-trial)

@@

s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg)

sum1 <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == .15))

k <- sum1 |> group_by(id) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
  mutate(matched_goal = (pct_change == .15))

outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>2.75)
outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()
```

@@

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  #facet_wrap(~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

# descriptive table to do with plot
sum1 |> 
  group_by(refClass,state) |> 
  summarise(mean_pct_change=median(pct_change), sd_pct_change=sd(pct_change), n=n()) |> 
  gt() |> 
  tab_header(title="Descriptive Statistics of Percent Change by State and Reference Class")

sum1 <- sum1 |> filter(!id %in% outlier_id)

@@

  # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
sum1 |> 
  filter(block==1) |>
  ungroup() |> 
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=.15),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~refClass,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)))
```

@@

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- sum1 |>
  group_by(refClass) |>
  summarise(
    matched_count = sum(matched_goal),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

@@

@@

>>>>

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,brms,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,ggh4x,lme4,flextable,kableExtra)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())
```

@@

s1 <- s1 |> group_by(id) |> 
  mutate(trial=row_number(), block=ifelse(trial>=12,2,1), plan=as_factor(plan),pct_goal=.15) |> 
  select(-trial)

@@

s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator"))

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

k <- s1_agg |> group_by(id,pct_goal) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

s1_agg2 <- s1_agg |> group_by(id,refClass,state) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

outliers <- s1_agg |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>3.0)
outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()

```{r}
#| panel: tabset
#| output: asis

t1 <- datasummary_crosstab(data=s1_agg |> group_by(id) |> slice(1),
                           calc~refClass,
                           statistic = 1 ~ N + Percent("row"),
                           output='markdown')
t2 <- datasummary(data=s1_agg4 ,(` `= pct)*refClass ~ Mean+Median+Mode+sd+Histogram,output='markdown',title="% of trials where goal was matched")
t3 <- datasummary(data=s1_agg, pct_change*refClass~Mean+Median+Mode, output='markdown')
list_of_tables <- list(t1,t2,t3)
names(list_of_tables) <- c("Calculator by Ref Class tallies","% of trials with matched goal","% change in plan")

for (itable in names(list_of_tables)) {
  cat(sprintf("\n## %s\n\n", itable))
  print(list_of_tables[[itable]])
}

@@

plot_hist_condit <- function(fcondit,dat,pvar) {
  full_range <- dat |> pull({{pvar}}) |> range(na.rm=TRUE)
  dat |>
    filter(refClass == fcondit) |> 
    ggplot(aes(.data[[pvar]])) +
    geom_bar(fill = 'dodgerblue4',width=1) +
    theme_minimal() #+ coord_cartesian(xlim = full_range)
}

s1_agg4 |> group_by(refClass) |> 
  summarise(Mean=mean(pct), Median=median(pct),Mode=modeest::mlv(pct,method='mfv'), sd=sd(pct), n=n()) |> 
  mutate(Distribution=refClass) |> rename("Reference Class"=refClass) |>  
  gt() |> 
  tab_header(title="Descriptive Statistics of % of trials with matched goal") |> 
  text_transform(
    locations=cells_body(columns='Distribution'),
    fn=function(column) {
      map(column,plot_hist_condit,dat=s1_agg4,pvar='mg') |> 
        ggplot_image(height=px(50),aspect_ratio=2)
    }
  )

m_zib <- brm(
  bf(pct ~ refClass + (1|id),
     zi ~ refClass + (1|id)),
  family = zero_inflated_beta(),
  data = s1_agg4
)

m_bb <- brm(
  pct ~ refClass + (1|id),
  family = beta_binomial(),
  data = s1_agg4
)

m_bb <- brm(
    successes | trials(4) ~ refClass,
    family = beta_binomial(),
    data = s1_agg4 |> 
        mutate(successes = pct * 4)  # convert proportion back to count
)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -0.52      0.01    -0.55    -0.49 1.00     4003     3155
# refClassPercentage    -0.73      0.02    -0.77    -0.69 1.00     3603     2648
# refClassUSD           -1.74      0.02    -1.79    -1.69 1.00     3478     3007
# 
# Further Distributional Parameters:
#     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# phi     0.33      0.00     0.33     0.34 1.00     3255     2429

# > m_bb |> emmeans(~refClass, type="response")
#  refClass   prob lower.HPD
#  upper.HPD
#  kWh        0.37      0.37      0.38
#  Percentage 0.22      0.22      0.23
#  USD        0.09      0.09      0.10

m_bb |> emmeans(~refClass)
m_bb |> emmeans(~refClass, type="response")

m_bb_re <- brm(
    successes | trials(4) ~ refClass + (1|id),
    family = beta_binomial(),
    cores=4,
    data = s1_agg4 |> 
        mutate(successes = pct * 4)
)

m_bb2 <- brm(
    mg | trials(4) ~ 0+refClass,
    family = beta_binomial(link = "identity", link_phi = "identity"),
    chains=2,
    cores=2,
    data = s1_agg4
)

m_bb3 <- brm(
    mg | trials(2) ~ 0+refClass,
    family = beta_binomial(link = "identity", link_phi = "identity"),
    chains=2,
    cores=2,
    data = s1_agg2
)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# refClasskWh            0.38      0.00     0.38     0.38 1.00     1900     1385
# refClassPercentage     0.22      0.00     0.22     0.22 1.00     1868     1549
# refClassUSD            0.09      0.00     0.09     0.10 1.00     2608     1328
# 
# Further Distributional Parameters:
#     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# phi     0.17      0.00     0.17     0.18 1.00      813      940

m_bb4 <- brm(
    mg | trials(2) ~ refClass + (1|id),
    family = beta_binomial(),
    chains=2,
    cores=2,
    data = s1_agg2
)

m_t1 <- brm(matched_goal ~ refClass + (1|id), family = bernoulli(), data = s1_agg)
 m_t1 |> emmeans(~refClass, type="response")

# 
# m_zib <- brm(
#   bf(pct ~ refClass + (1|id),
#      zi ~ refClass + (1|id)),
#   family = zero_inflated_beta(),
#   iter=2000,
#   cores=4,
#   data = s1_agg4 |> mutate( pct = ifelse(pct == 1, 0.999, ifelse(pct == 0, 0.001, pct)))
# )
# Multilevel Hyperparameters:
# ~id (Number of levels: 242) 
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# sd(Intercept)        0.09      0.07     0.00     0.26 1.00     2164     1682
# sd(zi_Intercept)     0.77      0.63     0.03     2.32 1.00     2846     1826
# 
# Regression Coefficients:
#                       Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept                -0.37      0.16    -0.67    -0.06 1.00     5376     2635
# zi_Intercept             -8.17      3.46   -17.15    -3.80 1.00     2673     1375
# refClassPercentage       -0.43      0.22    -0.87     0.00 1.00     5054     2984
# refClassUSD              -0.69      0.21    -1.10    -0.28 1.00     4759     2772
# zi_refClassPercentage    -0.04      4.78   -10.32    10.04 1.00     2037     1627
# zi_refClassUSD            0.47      4.50    -8.53    10.19 1.00     2181     1665

 s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
   group_by(refClass) |> 
  summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n())

 post <- posterior_samples(m_t1)
nsims <- 1000
re_sims <- rnorm(nsims, 0, mean(post$sd_id__Intercept))
```

```{r}
#| eval: false
s1_agg |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  #facet_wrap(~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

# descriptive table to do with plot
s1_agg |> 
  group_by(refClass,state) |> 
  summarise(mean_pct_change=median(pct_change), sd_pct_change=sd(pct_change), n=n()) |> 
  gt() |> 
  tab_header(title="Descriptive Statistics of Percent Change by State and Reference Class")

s1_agg <- s1_agg |> filter(!id %in% outlier_id)

@@

  # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
s1_agg |> 
  filter(block==1) |>
  ungroup() |> 
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=.15),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~refClass,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)))
```

@@

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- s1_agg |>
  group_by(refClass) |>
  summarise(
    matched_count = sum(matched_goal),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

@@

s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=pct)) +
  geom_density(aes(fill=refClass), alpha=0.5)

# histogram
s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=pct)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distributions of % of trials that matched goal",
       x="Percent of trials that matched goal",
       y="Number of Participants",
       fill="Reference Class") +
  scale_fill_brewer(palette="Set2") + 
  scale_y_continuous(breaks=seq(0,80,5)) + 
  scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))

# s1_agg |> group_by(id,refClass) |> 
#   summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
#   group_by(refClass) |>
#   summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n())

s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal), n=n(), pct=mg/n) |> 
  group_by(refClass,mg) |>
  summarise(n=n()) |>
  setNames(c("refClass", "# of trials with matched goal", "n")) |>
  knitr::kable(booktabs=T,col.names = c("Reference Class", 
                            "# of correct trials ",
                            "n"),
               format = "html") |> 
  kableExtra::kable_styling(full_width = F) |> column_spec(1:3, width = "2cm")

@@

contingency_table <- s1_agg %>%
    group_by(refClass) %>%
    summarise(
      matched = sum(matched_goal),
      unmatched = n() - sum(matched_goal)
    ) %>%
    select(matched, unmatched) %>%
    as.matrix()

# Perform the chi-squared test
chisq_test <- chisq.test(contingency_table)

# View the results
chisq_test

s1_aggb <- s1_agg %>%
  mutate(matched_goal = as.numeric(matched_goal))

m1_s1 <- brms::brm(
  matched_goal ~ refClass,        # Logistic regression with reference class as predictor
  data = s1_agg,                    # The dataset
  family = bernoulli(),           # Binary outcome model
 # prior = set_prior("normal(0, 1)", class = "b"),  # Prior for coefficients
  iter = 2000,                    # Number of iterations
  chains = 4,                     # Number of MCMC chains
  cores = 4,                      # Number of cores for parallel computation
  seed = 123                      # Set seed for reproducibility
)
# Regression Coefficients: - prior
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -0.53      0.12    -0.76    -0.30 1.00     3182     2941
# refClassPercentage    -0.76      0.18    -1.12    -0.41 1.00     2565     2685
# refClassUSD           -1.64      0.21    -2.06    -1.24 1.00     2576     2941
# Regression Coefficients: - no prior
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -0.49      0.12    -0.72    -0.27 1.00     3259     2975
# refClassPercentage    -0.81      0.19    -1.18    -0.45 1.00     2364     2585
# refClassUSD           -1.73      0.21    -2.15    -1.32 1.00     2643     2942

m2_s1 <- brms::brm(
  matched_goal ~ refClass + (1|id),       
  data = s1_agg,                    
  family = bernoulli(),           
  #prior = set_prior("normal(0, 1)", class = "b"),  
  iter = 2000,                    
  chains = 4,                     
  cores = 4,                      
  seed = 123                      
)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -3.97      0.95    -6.07    -2.38 1.00     1001     1715
# refClassPercentage    -0.54      0.75    -1.97     0.96 1.00     1523     2223
# refClassUSD           -2.26      0.76    -3.74    -0.77 1.00     1729     2581
# Regression Coefficients: - no prior
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -2.04      0.97    -4.11    -0.35 1.01      616     1384
# refClassPercentage    -3.16      1.35    -5.97    -0.65 1.00      542     1234
# refClassUSD           -6.84      1.68   -10.42    -3.87 1.01      581     1096

m3_s1 <- brms::brm(
  matched_goal ~ refClass + (1|id) + (1|state), 
  data = s1_agg,                    
  family = bernoulli(),           
  prior = set_prior("normal(0, 1)", class = "b"),  
  iter = 2000,                    #
  chains = 4,                     
  cores = 4,                      
  seed = 123                      
)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -4.08      1.05    -6.39    -2.20 1.00      651     1280
# refClassPercentage    -0.52      0.78    -2.05     1.02 1.01      769     1583
# refClassUSD           -2.25      0.78    -3.78    -0.72 1.00     1036     2002

m4_s1 <- brms::brm(
  matched_goal ~ refClass + (1+plan|id) + (1|state), 
  data = s1_agg,                    
  family = bernoulli(),           
  iter = 4000,                    #
  chains = 4,                     
  cores = 4,                      
  seed = 123                      
)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -4.24      1.12    -6.60    -2.18 1.00     1280      962
# refClassPercentage    -0.49      0.79    -1.99     1.12 1.00     2347     4012
# refClassUSD           -2.22      0.78    -3.74    -0.71 1.00     2395     4132

m5_s1 <- brms::brm(
  matched_goal ~ refClass + (1+plan|id) + (1|state), 
  data = s1_agg,                    
  family = bernoulli(),           
  iter = 4000,                    #
  chains = 4,                     
  cores = 4,                      
  seed = 123                      
)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -2.20      1.15    -4.56     0.02 1.00     1819     2875
# refClassPercentage    -3.34      1.53    -6.59    -0.64 1.00     1139     2377
# refClassUSD           -7.37      1.90   -11.64    -4.13 1.00     1706     3083

```
```{r}
#| eval: FALSE

stanplot(m5_s1)

mcmc_plot(m5_s1, pars = c("b_Intercept","refClassPercentage", "refClassUSD"), regex = FALSE)
mcmc_plot(m5_s1, regex_pars=c('b_*'), regex = FALSE)

m5_s1 |> emmeans(~refClass)
#  refClass   emmean lower.HPD upper.HPD
#  kWh          -2.2      -4.6       0.0
#  Percentage   -5.4      -8.5      -2.9
#  USD          -9.4     -13.4      -6.3
# 
# Point estimate displayed: median 
# Results are given on the logit (not the response) scale. 
# HPD interval probability: 0.95

m5_s1 |> emmeans(~refClass) |> gather_emmeans_draws() |> 
  ggplot(aes(x=refClass,y=.value,fill=refClass)) +
  stat_dist_pointinterval() + stat_halfeye(alpha=.2) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95))

# transform the log-odds to odds
